---
title: "Network analysis of youth climate change anxiety"
format: html
---

## Introduction

This document presents the results of a network analysis based on the code from a paper by Alexandre Heeren and colleagues on the psychology of climate change, looking at climate change anxiety in the general population and how it relates to worry. Here we are working with a dataset of youth data (ages 14-24) which was gathered in 2022 (?) from a Qualtrics panel. 

## Conceptual Replication of Heeren's Network Analysis 

First we will try to replicate the network analysis from the paper in the youth data. We don't have a measure for worry, but we do have psychological distress as measured by the DASS, which we will use in its place. 

```{r packages from the paper, include = FALSE, echo = FALSE, warning=FALSE}

#| echo: false
#------------PACKAGES ----------------
require(huge)
require(networktools)
require(qgraph)
#if (!require(parcor)) install.packages("parcor"); require(parcor)
require(Matrix)
require(psych)
require(dplyr)
require(ggplot2)
require(igraph)
require(car)
require(ppcor)
require(bootnet)
require(mgm)
require(corrplot)
require(digest)
require(devtools)
#if (!require(xlsx)) install.packages("xlsx"); require(xlsx)
require(qgraph)
require(bootnet)
require(foreign)
require(qgraph)
require(grDevices)
require(ape)
require(networktools)
require(matrixcalc)
require(psy)
require(psych)
require(ppcor)
require(bootnet)
require(bnlearn)
require(corrplot)
require(smacof)
require(colorspace)
require(eigenmodel)
require(network)

#if (!requireNamespace("BiocManager", quietly = TRUE))
#  install.packages("BiocManager")
#BiocManager::install("Rgraphviz")

require(tcltk)
require(latentnet)
require(rgl)

```


```{r import and wrangle data, echo = FALSE}
library(haven)
youth_climate_data <- read_sav("Youth Mental Health and Climate Change_2023_IMPUTED_FOR ANALYSIS_wID_877_01_02_2026.sav")

youth_climate_data$DASS_total <- (youth_climate_data$DASS_anxiety+youth_climate_data$DASS_depression+youth_climate_data$DASS_stress)
Cognitive_Emo <-  youth_climate_data$CCAS_CE_impairment
Functional <-  youth_climate_data$CCAS_functional_impair
Experience <-  youth_climate_data$CCAS_CC_experience
Pro_env_behav <-  youth_climate_data$CCAS_behav_engage
Distress <- youth_climate_data$DASS_total

DATA_combined <- as.data.frame(cbind(Cognitive_Emo,
                                     Functional,
                                     Pro_env_behav,
                                     Experience, Distress))

#summary(DATA_combined)
DATA_combined <- na.omit(DATA_combined) # Need to get rid of NAs to avoid errors later
table.s1<-data.frame()
if (!require(e1071)) install.packages("e1071"); require(e1071) # for skewness
# step 3 for each variable, computing the indices of interest 
for (i in 1:ncol(DATA_combined)){
  table.s1[i,1]<-colnames(DATA_combined)[i]
  table.s1[i,2]<-round(mean(DATA_combined[,i], na.rm=TRUE),2)
  table.s1[i,3]<-round(sd(DATA_combined[,i], na.rm=TRUE),2)
  table.s1[i,4]<-min(DATA_combined[,i], na.rm=TRUE)
  table.s1[i,5]<-max(DATA_combined[,i], na.rm=TRUE)
  table.s1[i,6]<-round(skewness(DATA_combined[,i], na.rm=TRUE),2)
  table.s1[i,7]<-round(kurtosis(DATA_combined[,i], na.rm=TRUE),2)
}
colnames(table.s1)<-c("variable", "M", "SD","Min", "Max", "Skewness", "Kurtosis")
table.s1 # look at the table.s1 in the console 



```

This plot reproduces Figure S1 from the paper 

```{r correlation plot }
#| echo: false
library(corrplot)
M <- cor(DATA_combined)
corrplot(M, method="shade") #Figure S1

check <- is.positive.definite(cor(DATA_combined))#true 

```

Let's check if the matrix is positive definite: 

```{r check if matrix is positive definite}
check 


```

Now let's check the matrix for bad pairs using the Goldbricker package. 

```{r check for bad pairs}
#| echo: false
## Use goldbricker function to search for potential "bad pairs"
## This function compares correlations in a psychometric network in order to identify
## nodes which most likely measure the same underlying construct (i.e., are colinear).
## A "bad pairs" includes two nodes that are highly correlated with each other and that 
## correlate with other nodes in the network in a highly similar pattern.
gb <- goldbricker(DATA_combined, p = 0.05, 
                  method = "hittner2003", threshold=0.25, 
                  corMin=.50, progressbar = TRUE)
## Produce list of "bad pairs"
gb

```

Well that's nice. Let's transform the data now. 

```{r transform data with nonparanormal transform}
#| echo: false

data_npn <- huge.npn(DATA_combined, npn.func = "shrinkage", npn.thresh = NULL, verbose = TRUE)
data_npn <- as.data.frame(apply(data_npn, use='pairwise.complete.obs', 2, as.numeric))
summary(data_npn)

```


Now let's do some fancy stuff and reproduce that network graph! 

```{r GGM & NODE PREDICTABILITY }
#| echo: false

group.ega<- c("Climate Anxiety", "Climate Anxiety", 
              "Climate Change", "Climate Change", "Distress") 
p <- ncol(data_npn)

data_npn <- as.matrix(data_npn) # DA: I had to add this line to avoid an error 

fit <- mgm(data_npn, 
           type = rep('g', p),
           lev = rep(1, p),
           rule.reg = 'OR')

pred <- predict(fit, data_npn, 
                errorCon = 'R2')
pred$error


```

OK, now let's make the graph!


```{r estimate network, echo = FALSE }
NET <- estimateNetwork(data_npn, default="ggmModSelect", tuning = 0) 
plot(NET,node.width=1.4,layout = 'spring', 
     label.prop = 0.9, borders = TRUE, 
     pie = pred$error[,2], pieColor = rep('#377EB8',p),
     labels = colnames(data_npn),
     edge.labels=FALSE, details=FALSE, groups = group.ega,color= c("grey", "grey", 
                                                                  "yellow", "yellow", "red"), 
     vsize=12, border.width=1.9, label.cex = .9,
     label.scale.equal=TRUE, 
     color='#bbbbbb', border.width=6, legend = FALSE)

centralityPlot(NET, theme_bw=FALSE, scale = "raw0", 
               include = c("ExpectedInfluence"), decreasing= TRUE)
```

These replicate Figures 1 and 2 in the paper. 

Now let's do the Lasso analysis 

```{r lasso analysis}
#| echo: false

Network <- EBICglasso(cor(data_npn), 
                      n = nrow(data_npn), 
                      gamma = 0.5, threshold = TRUE) 

y <- qgraph(Network, layout = 'spring', 
            details = FALSE, labels=colnames(data_npn),
            vsize=16, color='#bbbbbb', border.width=2, 
            edge.labels=FALSE, label.cex = 1.2, 
            label.scale.equal=TRUE, legend = FALSE)

centralityPlot(y, theme_bw=TRUE, 
               scale = "raw0", 
               include = c("ExpectedInfluence"), 
               decreasing= TRUE)
centrality_auto(y)

```

This reproduces Figures S2 and S3


```{r bridge centrality analysis, echo = FALSE}
adjm <- getWmat(NET)
matrix<- graph_from_adjacency_matrix(adjm, mode="upper", diag=FALSE, weighted=TRUE)


## "walktrap" Community Analysis
walktrap_community <- cluster_walktrap(matrix)
# head(walktrap_community)
plot_dendrogram(walktrap_community, mode="phylo")

```

This figure isn't anywhere in the paper, but it's in the R code so I thought I'd include it. 


```{r bridge estimation, echo = FALSE, warning=FALSE}

#| caption: Bridge expected influence of the Gaussian graphic model
## bridge estimation using the communites resulting from the spinglass algorithm
my_communities <- walktrap_community$membership
checkVector <- is.vector(my_communities)
b <-bridge(matrix,communities=my_communities)
BPLOT <- plot(b, 
              include=c("Bridge Expected Influence (1-step)"),
              theme_bw=FALSE, raw0 = TRUE, signed=TRUE, decreasing= TRUE)
BPLOT + theme(axis.text=element_text(size=6.5))

```

This replicates Figure 3 from the paper

```{r stability and accuracy of the GGM }

# This takes ages to run so leave it out for the quick demo version! 
nonParametricBoot <- bootnet(data_npn, nBoots = 1000, 
                             default = "ggmModSelect", 
                             type="nonparametric", 
                             statistics="all", communities=my_communities)
plot(nonParametricBoot, labels=FALSE, order="sample")
plot(nonParametricBoot, "edge", plot = "difference", onlyNonZero = TRUE, order = "sample")
summary(nonParametricBoot) %>% ungroup %>% filter(type == "edge") %>% arrange(-sample) %>% View
plot(nonParametricBoot, statistics="ExpectedInfluence", plot="difference",order = "mean")
plot(nonParametricBoot, statistics="bridgeExpectedInfluence", plot="difference",order = "mean")


```



```{r stability, echo = FALSE}

#stability of the expected influence centrality - this also takes ages

caseDroppingBoot <- bootnet(data_npn, nBoots = 1000, 
                            default = "EBICglasso",  type="case", 
                            statistics="all", communities=my_communities)
corStability(caseDroppingBoot)
plot(caseDroppingBoot, statistics = c("ExpectedInfluence","bridgeExpectedInfluence"))

```



```{r check restricted range problem, echo = FALSE }
### DIFFERENTIAL VARIABILITY OF THE GGM
#  restricted range problem
c<- centrality_auto(matrix)
x <- c$node.centrality$ExpectedInfluence 
SD <- apply(data_npn,2,sd)
cor(SD,x)
cor.test(SD,x) 

```

Now we do a quick Bayesian network analysis based on 50 random re-starts with 100 perturbations for each re-start. 

```{r bayesian network analysis}
# AIM 2: Estimate Bayesian network -----------------------
## Fit a first Bayesian network, based on 50 random re-starts and 100 perturbations for each re-start.
netdata <- as.data.frame(data_npn) # DA: Need to turn it back into a data frame for this analsyis
set.seed(123)
fitBN1 <- hc(netdata, restart = 50, perturb = 100)  ## hc gives directed graph
# fitBN1

bnlearn::score(fitBN1, data = netdata)         
astr <- arc.strength(fitBN1, netdata, "bic-g")  ## connection strength
astr[order(astr[,3]), ]  ## sorted edge strength from strongest to weakest

strength.plot(fitBN1, astr, shape = "ellipse")


```


Now we stabilize the network across multiple samples through bootstrapping:


```{r stabilise network }
## Now we stabilize the network across multiple samples through bootstrapping:
## Learn 10000 network structures (may take a few min)

# DA note: This was originally 012 as the seed - this produces a completely different result 
## to the paper, with "worry" being outside the complex with no arrows. 
## However, setting the seed as 321 produced an identical result to 123. Go figure. 
set.seed(321)
bootnet <- boot.strength(netdata, R = 10000, algorithm = "hc", 
                         algorithm.args = list(restart = 5, perturb = 10), debug = TRUE)  
head(bootnet)

# Save the bootstrapped networks
#saveRDS(bootnet, "./bootDAG.rds")

# Reload the saved bootstrapped networks
#bootnet <- readRDS("./bootDAG.rds")

## net#1 optimal cutpoint, according to Scurati & Nagarajan (2013)
avgnet <- averaged.network(bootnet)
avgnet      
bnlearn::score(avgnet, data = netdata) # -5126.457
thresh <- avgnet$learning$args[[1]]
thresh                                  ## optimal significance threshold = 0.3319
astr <- arc.strength(avgnet, netdata, "bic-g")   ## compute edge strengths
astr

suppressWarnings(strength.plot(avgnet, astr, shape = "ellipse"))
# FOR interpretation: Edge thickness signifies the probability of prediction is in the direction depicted.





```

